---
title: "Untitled"
output: html_document
date: '2023-10-10'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readxl) 
library(janitor)
library(viridis)
library(scales)
library(ggrepel)
library(plotly)
library(flextable)


theme_set(theme_light())

# Hopefully this allows the DT to show up
options(htmltools.preserve.raw = FALSE)

# disabling scientific notation
options(scipen = 100)

`%out%` <- Negate(`%in%`)

# function for transposing df
transpose_df <- function(df) {
  t_df <- data.table::transpose(df)
  colnames(t_df) <- rownames(df)
  rownames(t_df) <- colnames(df)
  t_df <- t_df %>%
    tibble::rownames_to_column(.data = .) %>%
    tibble::as_tibble(.)
  return(t_df)
}

# scaling functions 
range01 <- function(x){(x-min(x))/(max(x)-min(x))}
range_wna <- function(x){(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))}

#mode function 
mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

not_all_na <- function(x) any(!is.na(x))
not_any_na <- function(x) all(!is.na(x))



```

```{r message = FALSE, warning=FALSE}
locations <- read_xlsx("./data/ken_adminboundaries_tabulardata.xlsx", 
                      sheet = "ADM1") %>% 
  clean_names()

pcode1_shape <- 
  sf::st_read("./data/ken_adm_iebc_20191031_shp/ken_admbnda_adm1_iebc_20191031.shp", 
          quiet = TRUE) %>% 
  clean_names()

counties <- read_csv("./data/counties.csv") %>% 
  mutate(sex_modifier = fct_relevel(sex_modifier, c("male", "female", "total")), 
         county = recode(county, "national" = "National"))

kcse <- read_csv("./data/kcse.csv") %>% 
  mutate(no_mg = ifelse(mg %in% c("P", "U", "X", "Y"), 
                        TRUE, FALSE), 
         subject = ifelse(subject == "biology_for_the_blind", "biology", subject),
         area = recode(area, 
                       "kk" = "kakuma")) %>% 
  mutate(is_stem = ifelse(subject %in% c("biology", "chemistry", "computer_studies", 
                                         "general_science", "maths", "physics"), 
                          "STEM", "non-STEM"))

dadaab <- kcse %>% filter(area == "dadaab")

score_convert<- tribble(~grade, ~score, 
        "A", 12, 
        "A-", 11, 
        "B+", 10,
        "B", 9,
        "B-", 8,
        "C+", 7,
        "C", 6,
        "C-", 5,
        "D+", 4,
        "D", 3,
        "D-", 2,
        "E", 1,
        "X", 0,
        "U", 0, 
        "Y", 0, 
        "P", 0, 
        "W", 0)
```


```{r eval=FALSE}
kcse %>% 
  select(id,
         sex,  
         school, 
         subject,
         subject_group, 
         grade, 
         year, 
         area) %>% 
  write_csv("./data/example_template_anonymised.csv")

```



# Summary


```{r}
summary_flat <- kcse %>% 
  mutate(area = ifelse(area != "dadaab", "Kakuma_Kalobeyei", "Dadaab")) %>% 
  filter(subject %out% c("maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  group_by(year, area, subject, is_stem) %>% 
  summarise(score = mean(score, na.rm = TRUE)) %>% 
  pivot_wider(names_from = year, values_from = score, names_prefix = "year_") %>% 
  filter(!is.na(year_2018) & !is.na(year_2022) & year_2018 != 0) %>%
  mutate(increase = (year_2022 - year_2018) / year_2018)

summary_flat %>% 
  ggplot(aes(x =  is_stem, y = increase, colour = is_stem)) + 
  geom_boxplot(size = 1) +
  geom_point(size = 2, alpha = .5) +
  scale_colour_manual(values = c("darkgoldenrod3", "darkslategray4"), 
                      breaks = c("STEM", "non-STEM")) +
  theme(legend.position = "none") +
  scale_y_continuous(labels = percent) +
  labs(x = "", 
       y = "Increase in scores between 2018 and 2020")

t.test(increase ~ is_stem, data = summary_flat, paired = FALSE)

summary_flat %>% arrange(desc(increase))
```



Much poorer results in Kakuma and Kalobeyei.


```{r}
kcse %>%
  filter(area != "dadaab") %>% 
  mutate(is_stem = ifelse(subject %in% c("biology", "chemistry", "computer_studies", 
                                         "general_science", "maths", "physics"), 
                          TRUE, FALSE)) %>% 
  group_by(subject, year, is_stem) %>% 
  summarise(score = mean(score, na.rm = TRUE)) %>% 
  filter(subject %out% c("biology_for_the_blind", 
                         "maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  group_by(subject, is_stem) %>% 
  mutate(yoy_change = score - lag(score), 
         pc_change = yoy_change / score, 
         overall_change = score[year == 2022] - score[year == 2018],
         pc_overall_change = overall_change / score[year == 2018]) %>% 
  filter(!is.na(yoy_change)) %>% 
  summarise(yoy_change = mean(yoy_change), 
            pc_change = mean(pc_change), 
            overall_change = mean(overall_change), 
            pc_overall_change = mean(pc_overall_change)) %>%
  mutate(subject = str_replace_all(subject, "_", " ")) %>% 
  ggplot(aes(x = pc_overall_change, 
             y = fct_reorder(str_to_title(subject), pc_overall_change), 
             fill = is_stem)) + 
  geom_col() + 
  geom_text(aes(label = percent(pc_overall_change, accuracy = 1)), 
            hjust = "inward") + 
  scale_x_continuous(labels = percent) + 
  scale_fill_manual(values = c("darkgoldenrod3", "darkslategray4"), 
                      breaks = c(TRUE, FALSE)) +
  labs(x = "% change between 2018 and 2022", 
       y = "", 
       title = "% change in mean subject scores between 2018 and 2022", 
       subtitle = "In Kakuma and Kalobeyei", 
       fill = "STEM")
```






Do it by sex as well 

```{r}
summary_flat <- kcse %>% 
  mutate(area = ifelse(area != "dadaab", "Kakuma_Kalobeyei", "Dadaab")) %>% 
  filter(subject %out% c("maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  mutate(is_stem = ifelse(subject %in% c("biology", "chemistry", "computer_studies", 
                                         "general_science", "maths", "physics"), 
                          "STEM", "non-STEM")) %>% 
  group_by(year, area, subject, is_stem) %>% 
  summarise(score = mean(score, na.rm = TRUE)) %>% 
  pivot_wider(names_from = year, values_from = score, names_prefix = "year_") %>% 
  filter(!is.na(year_2018) & !is.na(year_2022) & year_2018 != 0) %>%
  mutate(increase = (year_2022 - year_2018) / year_2018)

summary_flat %>% 
  ggplot(aes(x =  is_stem, y = increase, colour = is_stem)) + 
  geom_boxplot(size = 1) +
  geom_point(size = 2, alpha = .5) +
  scale_colour_manual(values = c("darkgoldenrod3", "darkslategray4"), 
                      breaks = c("STEM", "non-STEM")) +
  theme(legend.position = "none") +
  scale_y_continuous(labels = percent) +
  labs(x = "", 
       y = "Increase in scores between 2018 and 2020")

t.test(increase ~ is_stem, data = summary_flat, paired = FALSE)

summary_flat %>% arrange(desc(increase))
```

It is likely that our sample size isn't large enough. 

Overall, scores improved more in Dadaab than they did in Kakuma and Kalobeyei. However, whilst STEM subjects made larger improvements in Dadaab, the opposite was true in Kakuma and Kalobeyei, where non-STEM subjects improved more. Kakuma and Kalobeyei saw decreases in their STEM and non-STEM scores. 

<br>

```{r}
summary_flat %>% 
  ggplot(aes(x =  area, y = increase, colour = area)) + 
  geom_boxplot(size = 1) +
  geom_point(size = 2, alpha = .5) +
  scale_colour_manual(values = c("salmon", "blue2")) +
  scale_y_continuous(labels = percent) +
  facet_wrap(~ is_stem) +
  labs(x = "", 
       y = "Increase in scores between 2018 and 2020") + 
   theme(legend.position = "none", 
         strip.background = element_rect(fill = "black"), 
         strip.text = element_text(colour = "white", face = "bold"), 
         axis.text.x = element_text(size = 10))
```

# Dadaab

## Candidates assessed 

```{r}
kcse %>% filter(area == "dadaab" & !is.na(sex)) %>% 
  filter(mg %out% c("P", "U", "X", "Y")) %>% 
  distinct(id, sex, year) %>% 
  group_by(year, sex) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = year, y = count, fill = sex)) + 
  geom_col(position = position_dodge()) + 
  geom_text(aes(label = comma(count)), 
            position = position_dodge(width = .9), 
            vjust = -.3) + 
  scale_fill_manual(values = c("#1fa187", "#440154")) + 
  labs(title = "KCSE Candidates assessed by year in Dadaab", 
       y = "Number of children", 
       x = "")
```


<br>

Maybe not necessary? 

```{r}
dadaab %>%
  distinct(id, sex, year, no_mg) %>% 
  group_by(no_mg, sex, year) %>%
  summarise(count = n()) %>% 
  group_by(year, sex) %>% 
  mutate(subtotal = sum(count), 
         pc = count / subtotal) %>%
  filter(no_mg == TRUE) %>% 
  ggplot(aes(x = year, y = pc, fill = sex)) + 
  geom_col(position = position_dodge()) + 
  geom_text(aes(label = percent(pc, accuracy = .1)), 
            position = position_dodge(width = .9), 
            vjust = -.3) + 
  scale_fill_manual(values = c("#1fa187", "#440154")) + 
  scale_y_continuous(labels = percent) + 
  labs(title = "Percentage of KCSE candidates in Dadaab",
       subtitle = "who were absent or have ungraded or pending grades",
       y = "Percentage of candidates", 
       x = "")


```

## Mean scores

Overall, there has been an increase in the mean scores (and consequently AGPs) of KCSE candidates of both sexes, though males do outperform females, at least in the observed population. 

<br>



```{r}

kcse %>% filter(area == "dadaab" & !is.na(sex)) %>% 
  group_by(year, sex) %>% 
  summarise(mean_score = mean(mean_score, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = mean_score, fill = sex)) +
  geom_col(position = position_dodge()) + 
  geom_text(aes(label = round(mean_score, digits = 2)), 
            position = position_dodge(width = .9), 
            vjust = -.3) + 
  scale_fill_manual(values = c("#1fa187", "#440154")) + 
  labs(title = "Mean score by sex and year", 
       y = "Number of children", 
       x = "")


```

<br>


```{r}
kcse %>% filter(area == "dadaab" & !is.na(sex)) %>% 
  filter(agp > 0) %>% 
  group_by(year, sex) %>% 
  summarise(mean_agp = mean(agp, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = mean_agp, colour = sex)) + 
  geom_line(size = 1) + 
  geom_text(aes(label = round(mean_agp, digits = 2)), 
            vjust = -.3, hjust = "inward") + 
  scale_colour_manual(values = c("#1fa187", "#440154")) +
  facet_wrap(~ sex) + 
  labs(title = "Improvements in mean AGP across both sexes in Dadaab") +
  theme(legend.position = "none", 
        strip.background = element_rect(fill = "black"), 
        strip.text = element_text(face = "bold", size = 8))
```


## Distribution of scores by sex


The plot below shows the distributions of average STEM score of male and female candidates by year. Of note -- years 2018 and 2020 have markedly lower mean STEM scores across both males and female candidates. 

<br>

```{r}

dadaab_mean_stem <- kcse %>% 
  filter(area == "dadaab" & !is.na(sex) & is_stem == "STEM" & year != 2019) %>%
  group_by(year, sex) %>% 
  summarise(mean_stem_score = mean(score, na.rm = TRUE))

kcse %>% 
  filter(area == "dadaab" & !is.na(sex) & is_stem == "STEM" & year != 2019) %>%
  ggplot(aes(x = score, colour = sex)) + 
  geom_density(size = .4, alpha = .3, adjust = 3) + 
  scale_colour_manual(values = c("#11b86c", "#a90ecf")) +
  geom_vline(data = dadaab_mean_stem, aes(xintercept = mean_stem_score, 
                                          colour = sex), 
             lty = 2, size = .1, alpha = .8, show.legend = FALSE) +
  scale_x_continuous(breaks = seq(0, 12, by = 1)) +
  facet_wrap(~ year) +  
  theme(strip.background = element_rect(fill = "black"), 
        strip.text = element_text(face = "bold", size = 8), 
        plot.caption = element_text(hjust = .5)) +
   guides(colour = guide_legend(override.aes = list(shape = 19))) + 
  labs(x = "Candidate STEM scores", 
       y = "Density", 
       title = "Distribution of mean STEM scores by year and sex", 
       subtitle = "Mean score by sex marked by dotted lines", 
       colour = "Sex", 
       caption = "2019 excluded due to small sample size")

  
```

<br>

Additionally, whilst male candidates have outperformed female candidates in STEM subjects consistently, the difference in the results of the two sexes starting becoming more prominent in 2021, with the gap widening further in 2022: 

<br> 

```{r}
kcse %>% 
  filter(area == "dadaab" & !is.na(sex) & is_stem == "STEM" & year != 2019) %>%
  group_by(year, sex) %>% 
  summarise(mean_stem_score = mean(score, na.rm = TRUE)) %>% 
  pivot_wider(names_from = sex, values_from = mean_stem_score) %>% 
  rename(Year = year) %>% 
  mutate(Year = factor(Year)) %>% 
  mutate_at(vars("Female", "Male"), ~ round(., digits = 3)) %>% 
  flextable() %>% 
  theme_zebra() %>% 
  set_table_properties(layout = "autofit", width = .4) %>% 
  set_caption("Mean STEM scores in Dadaab")
```


In Dadaab, maths_b, general science 

Not really useful

```{r}

dadaab_subtotals <- dadaab %>% 
  filter(grade %out% c("X", "Y", "U", "P", "--")) %>% 
  group_by(year) %>% 
  summarise(count = n())

dadaab %>% 
  filter(grade %out% c("X", "Y", "U", "P", "--")) %>% 
  group_by(subject, year) %>%
  summarise(count = n()) %>% 
  filter(subject %out% c("biology_for_the_blind", 
                         "maths_b", 
                         "general_science")) %>% 
  ggplot(aes(x = year, y = count, fill = subject))  + 
  geom_col() + 
  geom_text(aes(year, count, label = comma(count), fill = NULL), 
            data = dadaab_subtotals, 
            vjust = -.5) + 
  scale_y_continuous(labels = comma) + 
  labs(title = "Number of papers graded per year per subject in Dadaab")


```

<br>

The most common subjects taken were largely compulsory subjects such as English, Maths and Kiswahili. Though of the sciences, Chemistry and Biology were much more widely taken than Physics. 

```{r}
dadaab %>% 
  filter(subject %in% c("general_science", 
                        "maths_b")) %>% 
  count(year, school)
  
  count(subject)
```

Not useful either. Perhaps more popular subjects by school or area 

```{r}
dadaab %>% 
  group_by(subject_group, subject) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% 
  mutate(pc = round(count / sum(count) * 100, digits = 2)) %>% 
  mutate(subject = ifelse(pc < 0.5, "other", subject), 
         subject_group = ifelse(subject == "other", 
                                NA_character_, 
                                subject_group)) %>% 
  group_by(subject_group, subject) %>% 
  summarise(count = sum(count), 
            pc = sum(pc)) %>% 
  arrange(desc(pc)) %>% 
  rename(`%_total` = pc) %>%
  flextable() %>% 
  theme_zebra() %>% 
  set_caption("Number of papers graded from Dadaab, 2018-2022") %>% 
  footnote(i = 12, j = 2, ref_symbols = "1", 
           as_paragraph("Includes Maths B, General Science, Biology for the Blind and Kenyan Sign Language")) %>% 
  fontsize(size = 8, part = "footer")


```

<br>



## Trends in STEM subject scores

The plot below shows the mean increase in scores between 2018 and 2020 per subject: 

```{r}
kcse %>% 
  filter(area == "dadaab") %>%
  group_by(subject, year, is_stem) %>% 
  summarise(score = mean(score, na.rm = TRUE)) %>% 
  filter(subject %out% c("biology_for_the_blind", 
                         "maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  group_by(subject, is_stem) %>% 
  mutate(yoy_change = score - lag(score), 
         pc_change = yoy_change / score, 
         overall_change = score[year == 2022] - score[year == 2018],
         pc_overall_change = overall_change / score[year == 2018]) %>% 
  filter(!is.na(yoy_change)) %>% 
  summarise(yoy_change = mean(yoy_change), 
            pc_change = mean(pc_change), 
            overall_change = mean(overall_change), 
            pc_overall_change = mean(pc_overall_change)) %>%
  mutate(subject = str_replace_all(subject, "_", " ")) %>% 
  ggplot(aes(x = pc_overall_change, 
             y = fct_reorder(str_to_title(subject), pc_overall_change), 
             fill = is_stem)) + 
  geom_col() + 
  geom_text(aes(label = percent(pc_overall_change, accuracy = 1)), 
            hjust = "inward") + 
  scale_x_continuous(labels = percent) + 
  scale_fill_manual(values = c("darkgoldenrod3", "darkslategray4"), 
                      breaks = c("STEM", "non-STEM")) +
  labs(x = "% change between 2018 and 2022", 
       y = "", 
       title = "% change in mean subject scores between 2018 and 2022", 
       subtitle = "In Dadaab", 
       fill = "STEM")

```

<br>

And this plot presents a more detailed view of the increases. 

> Score distribution geom_density per school. 

This plot shows the density, so that the distributions of the scores of male and female candidates can be more easily compared, even though there were many more male candidates than female ones. 

<br>


```{r fig.width=8, fig.height=6}
dadaab %>% 
  mutate(is_stem = ifelse(subject %in% c("biology", "chemistry", "computer_studies", 
                                         "general_science", "maths", "physics"), 
                          TRUE, FALSE)) %>% 
  group_by(subject, year, is_stem) %>% 
  summarise(score = mean(score, na.rm = TRUE)) %>% 
  mutate(score_round = round(score)) %>% 
  left_join(score_convert, 
            by = c("score_round" = "score")) %>% 
  filter(subject %out% c("maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  mutate(label = subject, 
         label = ifelse(year != 2021, "", subject), 
         subject = str_to_title(subject)) %>% 
  ggplot(aes(x = year, y = score, colour = is_stem)) + 
  geom_text_repel(aes(label = grade), size = 3, show.legend = FALSE) + 
  geom_line(size = 1) +
  facet_wrap(~ subject) +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10)) + 
  scale_colour_manual(values = c("darkgoldenrod3", "darkslategray4"), 
                      breaks = c(TRUE, FALSE)) +
  labs(title = "Average score per subject 2018-2022", 
       subtitle = "In Dadaab", 
       x = "", y = "Average score", 
       colour = "STEM subject") + 
  theme(legend.position = "bottom", 
        strip.background = element_rect(fill = "black"), 
        strip.text = element_text(face = "bold", size = 8), 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), 
        legend.box.margin = unit(c(-5, 0, 0, 0), "mm")) 
  
```

<br>

```{r}
dadaab_t_test <- dadaab %>% 
  filter(subject %out% c("maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  mutate(is_stem = ifelse(subject %in% c("biology", "chemistry", "computer_studies", 
                                         "general_science", "maths", "physics"), 
                          "STEM", "non-STEM")) %>% 
  group_by(year, subject, is_stem) %>% 
  summarise(score = mean(score, na.rm = TRUE)) %>% 
  pivot_wider(names_from = year, values_from = score, names_prefix = "year_") %>% 
  filter(!is.na(year_2018) & !is.na(year_2022) & year_2018 != 0) %>%
  mutate(increase = (year_2022 - year_2018) / year_2018) %>% 
  t.test(increase ~ is_stem, data = ., paired = FALSE) %>% 
  broom::tidy()

```

From the charts above, it does seem that the scores of STEM subjects have increased more than non-STEM subjects. Our hypothesis, then, is that UNICEF-assisted programmes, intervening in STEM education, have had a demonstrable impact on KCSE scores. That is why they are being compared to the increase in scores of non-STEM subjects. Let us take a look at the boxplot: 

```{r}
dadaab_flat <- kcse %>% 
  filter(!is.na(school) & area == "dadaab") %>% 
  filter(subject %out% c("maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  
  group_by(year, school, subject, is_stem) %>% 
  summarise(score = mean(score, na.rm = TRUE)) %>% 
  pivot_wider(names_from = year, values_from = score, names_prefix = "year_") %>% 
  filter(!is.na(year_2018) & !is.na(year_2022) & year_2018 != 0) %>%
  mutate(increase = (year_2022 - year_2018) / year_2018) 

dadaab_flat %>% 
  ggplot(aes(x =  is_stem, y = increase, colour = is_stem)) + 
  geom_boxplot(size = .7) +
  geom_point(size = 2, alpha = .3) +
  scale_colour_manual(values = c("darkgoldenrod3", "darkslategray4"), 
                      breaks = c("STEM", "non-STEM")) +
  scale_y_continuous(labels = percent) +
  labs(x = "", 
       y = "Increase in scores", 
       title = "Change in scores in Dadaab, 2018 - 2022") + 
  coord_cartesian(ylim = c(-1, 6)) + 
  theme(legend.position = "none", 
        strip.background = element_rect(fill = "black"), 
        strip.text = element_text(face = "bold"))

```
<br>

However, upon performing a t-test, we note that the p-value of `r dadaab_t_test %>% pull(p.value) %>% round(digits = 3)`when comparing the change in scores between STEM and non-STEM subjects between 2018 and 2022. This falls outside of the commonly-used threshold of <=0.05 used for statistical significance. The p.value tells us how likely a set of observations could have occurred by chance. 

The t-statistic, which tells us the size of the effect, is also quite small at `r dadaab_t_test %>% pull(statistic) %>% round(digits = 3)`. 

This doesn't necessarily rule out UNICEF-supported programmes having an effect on STEM scores. Let us first take a look at the increases in STEM and non-STEM scores at the school level: 

<br>

```{r}

kcse %>% 
  filter(area == "dadaab") %>% 
  filter(!is.na(class)) %>% 
  filter(subject %out% c("maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  group_by(year, school, sex, subject, is_stem) %>% 
  summarise(score = mean(score, na.rm = TRUE)) %>% 
  pivot_wider(names_from = year, values_from = score, names_prefix = "year_") %>% 
  filter(!is.na(year_2018) & !is.na(year_2022) & year_2018 != 0) %>%
  mutate(increase = (year_2022 - year_2018) / year_2018) %>% 
  ggplot(aes(x =  is_stem, y = increase, colour = is_stem)) + 
  geom_boxplot(size = .7) +
  geom_point(size = 2, alpha = .3) +
  scale_colour_manual(values = c("darkgoldenrod3", "darkslategray4"), 
                      breaks = c("STEM", "non-STEM")) +
  facet_wrap(~ school) + 
  scale_y_continuous(labels = percent) +
  labs(x = "", 
       y = "Increase in scores", 
       title = "Change in scores by school, 2018 - 2022") + 
  coord_cartesian(ylim = c(-1, 6)) + 
  theme(legend.position = "none", 
        strip.background = element_rect(fill = "black"), 
        strip.text = element_text(face = "bold"))

```

<br>
Scores overall in Dagahaley and Tawakal were largely flat. Though STEM subjects outperformed non-STEM subjects in 5 out of the 6 schools assessed in Dadaab. Notable improvements in all scores were noted In Hagadera, Ifo, Towfiq and Waberi (though non-STEM subjects increased more than STEM subjects in Waberi). Hagadera has also improved the most, with both STEM and non-STEM scores have increased by more than 200% between 2018 and 2020. 

Let us again perform t-tests on the school-level changes in STEM scores: 

* The `Estimate` indicates the difference between STEM and non-STEM subjects. If STEM subjects performed better overall, this number will be negative. The larger the number, the larger the difference between STEM and non-STEM scores. 

* The `T-statistic` is used to determine if there is a significant difference between the mean STEM and non-STEM scores. Larger absolute values indicate greater differences. 

* The `P-value` is the probability of the assumption that UNICEF-supported interventions have had no effect on STEM scores (null hypothesis) i.e. that differences in scores are due to chance. Lower values indicate that it is more likely that UNICEF-supported interventions have had a demonstrable effect. The commonly-accepted threshold for statistical significance is >=0.05. 

<br>

```{r}
kcse %>% 
  filter(area == "dadaab") %>% 
  filter(!is.na(school)) %>% 
  filter(subject %out% c("maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  group_by(year, school, subject, is_stem) %>% 
  summarise(score = mean(score, na.rm = TRUE)) %>% 
  pivot_wider(names_from = year, values_from = score, names_prefix = "year_") %>% 
  filter(!is.na(year_2018) & !is.na(year_2022) & year_2018 != 0) %>%
  mutate(increase = (year_2022 - year_2018) / year_2018) %>% 
  group_by(school) %>% 
  nest() %>% 
  mutate(t_test = map(data, ~ t.test(.x$increase ~ .x$is_stem, paired = FALSE)), 
         res = map(t_test, broom::tidy)) %>% 
  unnest(res) %>%  
  select(school, estimate, statistic, p.value) %>% 
  mutate_if(is.numeric, ~ round(., digits = 3)) %>% 
  arrange(estimate) %>% 
  rename(School = school, 
         Estimate = estimate, 
         `T-statistic` = statistic, 
         `P-value` = p.value) %>% 
  flextable() %>% 
  theme_zebra() %>% 
  set_caption("T-test results of score increases in STEM and non-STEM scores by school") %>% 
  set_table_properties(width = .6, layout = "autofit")
```

<br>

Neither the results at the area-level (Dadaab) or the school level are statistically significant. This does not necessarily rule out UNICEF-supported programmes having an effect on STEM scores -- the number of schools is fairly small, since one observation here is the change in mean score per subject per school across Dadaab as a whole.  

Additional research is warranted, especially to understand why STEM subjects performed so much better in Hagadera, Ifo and Towfiq. 

<br>

```{r}
kcse %>% 
  filter(area == "dadaab" & !is.na(school) & school != "NASIB") %>%
  filter(subject %out% c("maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  group_by(year, school, is_stem) %>% 
  summarise(score = mean(score, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = score, colour = is_stem)) + 
  geom_point(size = 1) + 
  geom_line(size = 1) + 
  geom_text_repel(aes(label = round(score, digits = 2)), 
                  size = 2.5) +
  scale_colour_manual(values = c("darkgoldenrod3", "darkslategray4"), 
                      breaks = c("STEM", "non-STEM")) +
  facet_wrap(~ school) + 
  theme(legend.position = "none", 
        strip.background = element_rect(fill = "black"), 
        strip.text = element_text(face = "bold")) + 
  labs(title = "STEM and non-STEM scores over time in Dadaab, 2018-2022", 
       x = "", 
       y = "Mean score") + 
  theme(axis.text.x = element_text(angle = 60, 
                                   vjust = .5))

```



<br>

```{r}


kk_flat <- kcse %>% 
  filter(area != "dadaab") %>% 
  filter(subject %out% c("maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  mutate(is_stem = ifelse(subject %in% c("biology", "chemistry", "computer_studies", 
                                         "general_science", "maths", "physics"), 
                          TRUE, FALSE)) %>% 
  group_by(year, subject, is_stem) %>% 
  summarise(score = mean(score, na.rm = TRUE)) %>% 
  pivot_wider(names_from = year, values_from = score, names_prefix = "year_") %>% 
  filter(!is.na(year_2018) & !is.na(year_2022) & year_2018 != 0) %>%
  mutate(increase = (year_2022 - year_2018) / year_2018)

# Too few points of comparison
# Only graphically represent the data 
kk_flat %>% 
  ggplot(aes(x =  is_stem, y = increase)) + 
  geom_boxplot()

dadaab_flat %>% 
  ggplot(aes(x =  is_stem, y = increase)) + 
  geom_boxplot()
  

t.test(increase ~ is_stem, data = dadaab_flat, paired = FALSE)

t.test(increase ~ is_stem, data = kk_flat, paired = FALSE)

```        


# Kakuma and Kalobeyei





## Sex 

Female candidates improved their STEM scores more than male candidates in Dadaab, though it should also be mentioned that the median improvement (marked by the thick line in the middle) was more than 100%. Outcomes were quite different in Kakuma and Kalobeyei -- with both the scores of STEM subjects of candidates of both sexes experiencing a decrease, with the scores of females decreasing more than those of males. 

<br>

```{r}
kcse %>% 
  filter(area == "dadaab" & !is.na(school) & school != "NASIB") %>%
  filter(subject %out% c("maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  group_by(year, school, is_stem) %>% 
  summarise(score = mean(score, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = score, colour = is_stem)) + 
  geom_point(size = 1) + 
  geom_line(size = 1) + 
  geom_text_repel(aes(label = round(score, digits = 2)), 
                  size = 2.5) +
  scale_colour_manual(values = c("darkgoldenrod3", "darkslategray4"), 
                      breaks = c("STEM", "non-STEM")) +
  facet_wrap(~ school) + 
  theme(legend.position = "none", 
        strip.background = element_rect(fill = "black"), 
        strip.text = element_text(face = "bold")) + 
  labs(title = "STEM and non-STEM scores over time in Dadaab, 2018-2022", 
       x = "", 
       y = "Mean score") + 
  theme(axis.text.x = element_text(angle = 60, 
                                   vjust = .5))
```


```{r}
sex_flat <- kcse %>% 
  mutate(area = ifelse(area != "dadaab", "Kakuma_Kalobeyei", "Dadaab")) %>% 
  filter(subject %out% c("maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  mutate(is_stem = ifelse(subject %in% c("biology", "chemistry", "computer_studies", 
                                         "general_science", "maths", "physics"), 
                          "STEM", "non-STEM")) %>% 
  group_by(year, area, subject, sex, is_stem) %>% 
  summarise(score = mean(score, na.rm = TRUE)) %>% 
  pivot_wider(names_from = year, values_from = score, names_prefix = "year_") %>% 
  filter(!is.na(year_2018) & !is.na(year_2022) & year_2018 != 0) %>%
  mutate(increase = (year_2022 - year_2018) / year_2018)

sex_flat %>% 
  filter(!is.na(sex) & is_stem == "STEM") %>% 
  ggplot(aes(x =  sex, y = increase, colour = sex)) + 
  geom_boxplot(size = 1) +
  geom_point(size = 2, alpha = .5) +
  scale_colour_manual(values = c("#1fa187", "#440154")) + 
  scale_y_continuous(labels = percent) +
  facet_wrap(~ area) +
  labs(x = "", 
       y = "Increase in scores between 2018 and 2020", 
       title = "Comparison between Female and Male Candidates, STEM subjects") + 
  theme(legend.position = "none", 
         strip.background = element_rect(fill = "black"), 
         strip.text = element_text(colour = "white", face = "bold"), 
         axis.text.x = element_text(size = 10))



t.test(increase ~ is_stem, data = summary_flat, paired = FALSE)
```



## Notes

STEM subjects Chemistry, Biology, Physics, Mathematics and Computer studies 

Correlation between English and Kiswahili and performance in STEM subjects 

National results 

Performance by subject, by year, by sex
Mean grade by year, by sex at the national level 
sex ratios of candidates by subject and overall 

Prospects 
relook with the partner on information management

```{r}
kcse %>% glimpse()
```


```{r}
dadaab %>% 
  count(school)

kcse %>% 
  filter(area != "dadaab") %>% 
  count(school)

# MSS - life works tumaini


```





```{r}
dadaab_agp <- dadaab %>% 
  filter(mg %out% c("X", "Y", "U", "P", "--")) %>% 
  distinct(id, agp, year) %>% 
  group_by(year) %>% 
  summarise(agp = mean(agp, na.rm = TRUE))

dadaab %>% 
  filter(mg %out% c("X", "Y", "U", "P", "--")) %>% 
  distinct(id, agp, year) %>% 
  ggplot(aes(x = agp)) + 
  geom_histogram() + 
  scale_x_log10(breaks = c(1, 10, 30, 50, 100)) +
  facet_wrap(~year) + 
  geom_vline(aes(xintercept = agp, colour = factor(year)), data = dadaab_agp) + 
  geom_text(aes(label = round(agp, digits = 2)), data = dadaab_agp, 
            y = 200, hjust = -.15) + 
  labs("Average AGP by year") + 
  theme(legend.position = "none")

```


Candidates by subject, by year 

# Sex ratios

```{r}
dadaab %>% 
  filter(!is.na(sex)) %>% 
  group_by(sex) %>% 
  summarise(count = n_distinct(id)) %>% 
  mutate(subtotal = sum(count), 
         pc = count / subtotal) %>% 
  ggplot(aes(x = pc, y = sex, fill = sex)) + 
  geom_col() + 
  geom_text(aes(x = pc, 
                y = sex, 
                label = percent(pc, accuracy = 1)), 
            position = position_stack(vjust = .5), 
            colour = "white")  +
  scale_fill_manual(values = c("#35b779", "#443983")) +
  scale_x_continuous(labels = percent) + 
  labs(x = "", y = "", 
       title = "Sex ratios of assessed KCSE candidates", 
       subtitle = "In Dadaab County", 
       fill = "")
scale_colour_manual(values = c("darkgoldenrod3", "darkslategray4"), 
                      breaks = c(TRUE, FALSE)) +
  
```


```{r}
dadaab %>% 
  filter(!is.na(sex)) %>% 
  filter(subject %out% c("biology_for_the_blind", 
                         "maths_b", 
                         "general_science", 
                         "kenyan_sign_language")) %>% 
  mutate(label = paste0(subject_group, "_", str_to_title(subject))) %>% 
  group_by(label, sex) %>% 
  summarise(count = n()) %>% 
  mutate(subtotal = sum(count), 
         pc = count / subtotal) %>% 
  ggplot(aes(x = pc, y = fct_rev(label), fill = sex)) + 
  geom_col() + 
  geom_text(aes(x = pc, 
                y = label, 
                label = percent(pc, accuracy = 1)), 
            position = position_stack(vjust = .5), 
            colour = "white")  +
  scale_fill_manual(values = c("#35b779", "#443983")) +
  scale_x_continuous(labels = percent) + 
  labs(x = "", y = "", 
       title = "Sex ratios of assessed KCSE candidates", 
       subtitle = "In Dadaab County", 
       fill = "")
    

```



















